Fast Automatic Bayesian Cubature Using Lattice Sampling

### 2019-03-04 BayesianCubaturePaper2018 - Rev1.tex
Moved out full bayes derivations to Appendix and few other corrections

% This is a general template file for the LaTeX package SVJour3
% for Springer journals. Original by Springer Heidelberg, 2010/09/16
%
% Use it as the basis for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
\RequirePackage{amsmath}
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amssymb}
%
% insert here the call for the packages your document requires
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%\usepackage{latexsym}
% etc.
%
% Jag's 
\usepackage{tabu}
\usepackage{cancel}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
%\usepackage[caption=false]{subfig}
\usepackage{subcaption}
\usepackage{booktabs, mathtools}
\usepackage[dvipsnames]{xcolor}
\usepackage{url}
%\usepackage{refcheck}  % points out unused labels, refs ...

@@

Suppose that the integrand, $f$, is drawn from a Gaussian process, i.e., $f \sim \mathcal{GP}(m,s^2 C_\vtheta)$.  Specifically, $f$ has real-valued constant mean $m$ and covariance function $s^2C_\vtheta$, where $s$ is a non-negative scale factor, and $C_\vtheta: [0,1]^d \times [0,1]^d \to \mathbb{R} $ is a symmetric, positive-definite function and parameterized by $\vtheta$:
\begin{multline} \label{FJH:eq:CondPosDef}
\mC^T = \mC,  \quad \va^T \mC \va > 0, \quad \text{where }  \mC = \left(  C_\vtheta(\vx_i,\vx_j)  \right)_{i,j=1}^n,\\
 \text{for all } \va \ne 0, \;
 n\in \mathbb{N}, \\
 \vx_1, \ldots, \vx_n \in [0,1]^d, \; \text{s.t} \; \vx_i \neq \vx_j \forall i\neq j, \; 1 \leq i, j \leq n.
\end{multline}
The function, $C_{\vtheta}$, and the Gram matrix, $\mC$ depend implicitly on $\vtheta$, but the notation may omit this parameter for simplicity's sake.

@@

\subsubsection{Empirical Bayes}  \label{sec:MLE}
One approach is to estimate the parameters is via maximum likelihood estimation (MLE).  The log-likelihood function of the parameters given the function data $\vy$ is:
\begin{multline*}
l(s,m,\vtheta | \vy)
= -\frac{1}{2} s^{-2} (\vy-m\vone)^T\mCInv(\vy-m\vone) 
\\
 - \frac{1}{2} \log(\det\, \mC) - \frac{n}{2} \log(s^2) + \text{constants.}
\end{multline*}
Maximizing the log-likelihood first with respect to $m$, then with respect to $s$, and finally with respect to $\vtheta$ yields
\begin{align}
\label{eqn_m_MLE}
m_\MLE &= \frac{\vone^T \mCInv \vy }{ \vone^T \mCInv \vone}, \\
\nonumber
s^2_{\MLE}  
&= \frac{1}{n} (\vy-m_{\MLE}\vone)^T\mCInv(\vy-m_{\MLE}\vone) 
\\
\label{eqn_s2_MLE}
&= 
\frac{1}{n}
\vy^T 
\left[ 
\mCInv - 
\frac{ \mCInv \vone \vone^T \mCInv }{\vone^T\mCInv \vone}
\right] \vy, 
\end{align}
To obtain $\vtheta_\MLE$, substitute $m_\MLE$ and $s_\MLE$ into $l(s,m,\vtheta | \vy)$ and then numerically minimize
\begin{align}
\nonumber
\vtheta_\MLE
&= \argmin_{\vtheta} \biggl \{
\log\left(\vy^T 
\left[ \mCInv - 
\frac{ \mCInv \vone \vone^T \mCInv }{\vone^T\mCInv \vone}
\right] \vy 
\right)  \\
\label{eqn:thetaMLE}
 & \qquad +  \frac{1}{n} \log(\det(\mC))
\biggr \}.
\end{align}
The MLE estimate of $\vtheta$ balances minimizing the covariance scale factor, $s^2_{\MLE}$, against minimizing  $\det(\mC)$.

Under these estimates of the parameters, the cubature \eqref{eqn:BayesCub} and the credible interval \eqref{eqn_prob_confidence_interval} are explicitly written as
\begin{align} \label{eqn:cubMLE}
\hmu_\MLE  &= 
\left(
\frac{ (1 - \vone^T  \mCInv\vc )  \vone }{ \vone^T \mCInv \vone}   +  \vc 
\right)^T  \mCInv \vy, \\
\nonumber 
\err_\MLE^2 & : = \frac{2.58^2}{n}
 \vy^T \left[ \mCInv - 
\frac{ \mCInv \vone \vone^T \mCInv }{\vone^T\mCInv \vone}
\right] \vy \\
\label{eqn:errMLE}
&\qquad \qquad \times 
(c_0 - \vc^T\mC^{-1}\vc ),
\end{align}
\begin{align}
\label{eqn_prob_CI_MLE}
\mathbb{P}_f \left[
|\mu-\hmu_\MLE| \leq \err_\MLE \right]  = 99\%.
\end{align}
Here $c_0$, $\vc$, and $\mC$ are assumed implicitly to be based on $\vtheta = \vtheta_\MLE$.

\subsubsection{Full Bayes}

Rather than use maximum likelihood to determine $m$ and $s$ one can treat them as hyperparameters with a non-informative, conjugate prior, namely $\vrho_{m,s^2}(\xi, \lambda) \propto 1/\lambda$. Let us assume integral $\mu$, a random variable, taking typical values $z$, then the posterior density for the $\mu$ given data $\vy$ using Bayes theorem is
\begin{align*}
\MoveEqLeft[1]{\rho_{\mu}(z | \vf = \vy)} \\
& \propto \int_{0}^\infty \int_{-\infty}^\infty \rho_{\mu}(z | \vf = \vy, m = \xi, s^2 = \lambda) \\
& \hspace{2.5cm} \times  \rho_{\vf}(\vy | \xi, \lambda ) \rho_{m, s^2}(\xi, \lambda) \, \D \xi \D \lambda \\
& \propto \displaystyle \int_{0}^\infty  \frac{1}{\lambda^{(n+3)/2}} \\
& \times \int_{-\infty}^\infty  \exp \biggl( -\frac{1}{2\lambda}\biggl\{
\frac{
[z - \xi (1 - \vc^T \mC^{-1} \vone)  -  \vc^T \mC^{-1} \vy]^2}
{c_0  -\vc ^T \mC^{-1} \vc}  \\
& \hspace{1.8cm} + (\vy - \xi \vone)^T \mC^{-1}(\vy - \xi \vone) \biggr \} \biggr) \, \D \xi \D \lambda \\
& \hspace{2.5cm}
\text{by \eqref{eqn:fGaussDist}, \eqref{eqn:condInteg}} \; \text{and} \; \rho_{m,s^2}(\xi,\lambda) \propto 1/\lambda \\
& \propto \displaystyle \int_{0}^\infty  \frac{1}{\lambda^{(n+3)/2}} \cdots \\ 
& \qquad \cdots \int_{-\infty}^\infty  \exp\left( -\frac{\alpha \xi^2 -2 \beta \xi + \gamma}{2\lambda(c_0  -\vc ^T \mC^{-1} \vc)} \right) \, \D \xi \D \lambda, \\
\intertext{where}
\alpha & = (1 - \vc^T \mC^{-1} \vone)^2 + \vone^T \mC^{-1} \vone (c_0  -\vc ^T \mC^{-1} \vc),\\
\beta & =(1 - \vc^T \mC^{-1} \vone)(z - \vc^T \mC^{-1} \vy ) \\
& \qquad \qquad  + \vone^T \mC^{-1} \vy (c_0  -\vc ^T \mC^{-1} \vc),\\
\gamma &  = (z - \vc^T \mC^{-1} \vy )^2  + \vy^T \mC^{-1} \vy (c_0  -\vc ^T \mC^{-1} \vc).
\end{align*}
In the derivation above and below, factors that are independent of $\xi$, $\lambda$, or $z$ can be discarded since we only need to preserve the proportion.  But, factors that depend on $\xi$, $\lambda$, or $z$ must be kept.  
%Completing the square allows us to compute the integral with respect to $\xi$:
Completing the square, $
\alpha \xi^2 -2 \beta \xi + \gamma 
= \alpha (\xi -\beta/\alpha)^2  - (\beta^2/\alpha) + \gamma,
$
allows us to evaluate the integrals with respect to $\xi$ and $\lambda$:
\begin{align*}
\MoveEqLeft{\rho_{\mu}(z | \vf = \vy)} \\
& \propto \displaystyle \int_{0}^\infty  \frac{1}{\lambda^{(n+3)/2}}  \exp\left( -\frac{  \gamma - \beta^2/\alpha}{2\lambda(c_0  -\vc ^T \mC^{-1} \vc)} \right)  \cdots \\
& \qquad \qquad \cdots \int_{-\infty}^\infty  \exp\left( -\frac{\alpha (\xi -\beta/\alpha)^2}{2\lambda(c_0  -\vc ^T \mC^{-1} \vc)} \right) \, \D \xi \D \lambda \\
& \propto \displaystyle \int_{0}^\infty  \frac{1}{\lambda^{(n+2)/2}}  \exp\left( -\frac{  \gamma - \beta^2/\alpha}{2\lambda(c_0  -\vc ^T \mC^{-1} \vc)} \right) \D \lambda \\
& \propto \left(\gamma - \frac{\beta^2}{\alpha}\right)^{-n/2} \propto \left(\alpha \gamma - \beta^2\right)^{-n/2}.
\end{align*}
Finally, we simplify the key term via straightforward calculations to the following:
\begin{align*}
\alpha \gamma - \beta^2 \propto 1 +  \frac{(z - \hmu_{\textup{MLE}})^2}{(n-1)\widehat{\sigma}_{\textup{full}}^2}, \\
\end{align*}
where 
\begin{multline*}
%\label{FJH:eq:signmaFull}
\hsigma^2_{\textup{full}} 
:= \frac{1}{n-1}
\vy^T\left[ \mC^{-1} 
- \frac{ \mC^{-1} \vone\vone^T \mC^{-1}}{\vone^T \mC^{-1} \vone}  \right]\vy
\\ 
\times  \left[\frac{(1 - \vc^T \mC^{-1} \vone)^2}{\vone^T \mC^{-1} \vone} + (c_0  -\vc ^T \mC^{-1} \vc) \right].
\end{multline*}

@@

Because the shape parameter, $\vtheta$, enters the definition of the covariance kernel in a non-trivial way, the only way to treat it as a hyperparameter and assign a tractable prior would be for the prior to be discrete.  We believe in practice that choosing such a prior involves more guesswork than using the empirical Bayes estimate of $\vtheta$ in \eqref{eqn:thetaMLE} or the cross-validation approach described next.

@@

The loss function $\textup{GCV}$ depends on $m$ and $\vtheta$, but not on $s$.  Minimizing the GCV  yields
\begin{align*}
m_{\textup{GCV}} = \frac{\vone^T \mC^{-2} \vy}{\vone^T \mC^{-2} \vone},\\ 
\end{align*}
\begin{multline}
\label{vthetaGCV}
\vtheta_{\textup{GCV}} = \argmin_\vtheta \biggl\{\log \left(  \vy^T \left[\mC^{-2} - \frac{\mC^{-2} \vone \vone^T \mC^{-2}}{\vone^T \mC^{-2} \vone}  \right] \vy \right)  \\
- 2 \log \left ( \trace(\mC^{-1}) \right ) \biggr\} .
\end{multline}
Plugging this value of $m$ into \eqref{eqn:BayesCub} yields
\begin{align}
\label{eqn:muCV}
\widehat{\mu}_{\textup{GCV}}
= \left(\frac{(1 - \vone^T  \mC^{-1}\vc) \mC^{-1} \vone}{\vone^T \mC^{-2} \vone} + \vc \right)^T \mC^{-1} \vy.
\end{align}

@@

To demonstrate automatic Bayesian cubature consider a Mat\'ern covariance kernel:
\begin{align*}
C_{\theta}(\vx, \vt) = \prod_{k=1}^d \exp(-\theta|\vx_k-\vt_k|)(1+\theta |\vx_k-\vt_k|),
\end{align*}
and Sobol points as the nodes.
Also, consider the integration problem of evaluating  \emph{multivariate normal probabilities}:
\begin{align}
\label{eqn:GaussDef}
\mu = \int_{(\va,\vb)} \frac{\exp\bigl(- \frac 12 \vt^T \mSigma^{-1} \vt \bigr)}{\sqrt{(2 \pi)^d \det(\mSigma)}} \, \dvt,
\end{align}
where $(\va,\vb)$ is a finite, semi-infinite or infinite box in $\reals^d$.  This integral does not have an analytic expression for general $\mSigma$, so cubatures are required.

Genz \cite{Gen93} introduced a variable transformation to transform \eqref{eqn:GaussDef} into an integral on the unit cube.  Let $\mSigma= \mL \mL^T$ be the Cholesky decomposition where $\mL = (l_{jk})_{j,k=1}^d$ is a lower triangular matrix.  Iteratively define
\begin{align*}
\alpha_1& = \Phi(a_1), \qquad \beta_1 = \Phi(b_1)
\\
\alpha_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
a_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\\
\beta_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
b_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\end{align*}
\begin{align}
\label{fGenzdef}
f_{\text{Genz}}(\vx) = \prod_{j=1}^d [\beta_j(\vx) - \alpha_j(\vx)].
\end{align}
where $\Phi$ is the cumulative standard normal distribution function.  Then, $\mu = \int_{[0,1]^{d-1}} f_{\text{Genz}}(\vx) \, \dvx$.

\begin{figure}
	\captionsetup[subfigure]{labelformat=empty}
	%\begin{subfigure}[h]{0.48\linewidth}
	%	\includegraphics[width=1.1\linewidth]{Plotting_gaussian}
	%\end{subfigure}
	\centering
	%\begin{subfigure}[h]{0.48\linewidth}
		\includegraphics[width=0.7\linewidth]{GenzFunc_varTx_none}
	%\end{subfigure}
	\caption{The $d=3$ multivariate normal probability transformed to an integral of $f_{\text{Genz}}$ with  $d=2$. This plot can be reproduced using \code{IntegrandPlots.m} in GAIL}
	\label{fig:MVN_Genz}
\end{figure}

We use the following parameter values in the simulation: 
\begin{align*}
d = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol points \cite{DicEtal14a,DicPil10a}.  The results for  $\varepsilon = 10^{-2}, 10^{-3}, 10^{-4}$, and $10^{-5}$ and $100$ scrambles for each $\varepsilon$ are shown in \figref{fig:MVN_Metern_d2b2}.  We  observe the algorithm meets the error criterion 95\% of the time.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer, 
As shown in \figref{fig:MVN_Metern_d2b2}, computation time increases rapidly with $n$. 
The  maximum likelihood estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

@@

\subsection{Fast Transform Kernel}
We make some assumptions about the relationship between the covariance kernel and the nodes, which will be shown to hold in Section \ref{sec:shift_invariant_kernel} for rank-1 lattices and shift-invariant kernels.  First we introduce the notation
\begin{align}
\nonumber
\mC &= \Big(C_\vtheta(\vx_i,\vx_j)\Big)_{i,j=1}^n  = (\vC_1,...,\vC_n) 
\\
\label{eqn:ftk_factor}
&= \frac 1n \mV \mLambda \mV^H , 
\quad \quad \mV^H = n \mV^{-1}, \\
\nonumber
\mV &= (\vv_1,...,\vv_n)^T = (\vV_1,...,\vV_n), \\
\nonumber
\mC^p  &= \frac 1n \mV \mLambda^{p} \mV^H, \qquad \forall p \in \integers,
\end{align}
where $\mV^H$ is the Hermitian of $\mV$.  The columns of matrix $\mV$ are eigenvectors of $\mC$, and $\mLambda$ is a diagonal matrix of eigenvalues of $\mC$.
For any $n \times n$ vector $\vb$, define the notation  $\widetilde{\vb} := \mV^H \vb$.

@@

Under assumptions \eqref{fastcompAssump} the eigenvalues may be identified as the fast transform of the first column of $\mC$:
\begin{align}
\nonumber
\vlambda 
& = \begin{pmatrix}
\lambda_1 \\ \vdots \\ \lambda_n
\end{pmatrix} = \mLambda \vone = \mLambda \vv_1^* 
= \underbrace{\left( \frac 1n \mV^H  \mV \right) }_{\mathsf{I}} \mLambda \vv_1^* \\
&= \mV^H \left( \frac 1n \mV \mLambda \vv_1^* \right)
= \mV^H \vC_1 =  \widetilde{\vC}_1,
\label{eqn:fast_transform_to_eigvalues}
\end{align}
Where $\mathsf{I}$ is the identity matrix.
Also note that the fast transform of $\vone$ has a simple form
\begin{align*} 
\widetilde{\vone}
& = \mV^H \vone = \mV^H \vV_1 = \begin{pmatrix}n  \\ 0 \\ \vdots \\ 0 \end{pmatrix}.
\label{eqn:fast_transform_one}
\end{align*}

>>>>

% This is a general template file for the LaTeX package SVJour3
% for Springer journals. Original by Springer Heidelberg, 2010/09/16
%
% Use it as the basis for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
\RequirePackage{amsmath}
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amssymb}
%
% insert here the call for the packages your document requires
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%\usepackage{latexsym}
% etc.
%
% Jag's 
\usepackage{tabu}
\usepackage{cancel}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
%\usepackage[caption=false]{subfig}
\usepackage{subcaption}
\usepackage{booktabs, mathtools}
\usepackage[dvipsnames]{xcolor}
\usepackage{url}
\usepackage[titletoc,title]{appendix}
%\usepackage{refcheck}  % points out unused labels, refs ...

@@

Suppose that the integrand, $f$, is drawn from a Gaussian process (\cite{BriEtal18a}, \cite{Dia88a}, \cite{OHa91a}, \cite{Rit00a}, \cite{RasGha03a}), i.e., $f \sim \mathcal{GP}(m,s^2 C_\vtheta)$.  Specifically, $f$ has real-valued constant mean $m$ and covariance function $s^2C_\vtheta$, where $s$ is a non-negative scale factor, and $C_\vtheta: [0,1]^d \times [0,1]^d \to \mathbb{R} $ is a symmetric, positive-definite function and parameterized by $\vtheta$:
\begin{multline} \label{FJH:eq:CondPosDef}
\mC^T = \mC,  \quad \va^T \mC \va > 0, \quad \text{where }  \mC = \left(  C_\vtheta(\vx_i,\vx_j)  \right)_{i,j=1}^n,\\
 \text{for all } \va \ne 0, \;
 n\in \mathbb{N}, \qquad
 \vx_1, \ldots, \vx_n \in [0,1]^d, \; \text{s.t} \; \\\vx_i \neq \vx_j \forall i\neq j, \; 1 \leq i, j \leq n.
\end{multline}
The function, $C_{\vtheta}$, and the Gram matrix, $\mC$ depend implicitly on $\vtheta$, but the notation may omit this parameter for simplicity's sake.

@@

\subsubsection{Empirical Bayes}  \label{sec:MLE}
One approach is to estimate the parameters is via maximum likelihood estimation (MLE).  The log-likelihood function of the parameters given the function data $\vy$ is:
\begin{multline*}
l(s,m,\vtheta | \vy)
= -\frac{1}{2} s^{-2} (\vy-m\vone)^T\mCInv(\vy-m\vone) 
\\
 - \frac{1}{2} \log(\det\, \mC) - \frac{n}{2} \log(s^2) + \text{constants.}
\end{multline*}
Maximizing the log-likelihood first with respect to $m$, then with respect to $s$, and finally with respect to $\vtheta$ yields
\begin{align}
\label{eqn_m_MLE}
m_\MLE &= \frac{\vone^T \mCInv \vy }{ \vone^T \mCInv \vone}, 
\end{align}
\begin{align}
\nonumber
s^2_{\MLE}  
&= \frac{1}{n} (\vy-m_{\MLE}\vone)^T\mCInv(\vy-m_{\MLE}\vone) 
\\
\label{eqn_s2_MLE}
&= 
\frac{1}{n}
\vy^T 
\left[ 
\mCInv - 
\frac{ \mCInv \vone \vone^T \mCInv }{\vone^T\mCInv \vone}
\right] \vy, 
\end{align}
To obtain $\vtheta_\MLE$, substitute $m_\MLE$ and $s_\MLE$ into $l(s,m,\vtheta | \vy)$ and then numerically minimize
\begin{align}
\nonumber
\vtheta_\MLE
&= \argmin_{\vtheta} \biggl \{
\log\left(\vy^T 
\left[ \mCInv - 
\frac{ \mCInv \vone \vone^T \mCInv }{\vone^T\mCInv \vone}
\right] \vy 
\right)  \\
\label{eqn:thetaMLE}
 & \qquad +  \frac{1}{n} \log(\det(\mC))
\biggr \}.
\end{align}
The MLE estimate of $\vtheta$ balances minimizing the covariance scale factor, $s^2_{\MLE}$, against minimizing  $\det(\mC)$.

Under these estimates of the parameters, the cubature \eqref{eqn:BayesCub} and the credible interval \eqref{eqn_prob_confidence_interval} are explicitly written as
\begin{align} \label{eqn:cubMLE}
\hmu_\MLE  &= 
\left(
\frac{ (1 - \vone^T  \mCInv\vc )  \vone }{ \vone^T \mCInv \vone}   +  \vc 
\right)^T  \mCInv \vy,
\end{align}
\begin{align}
\nonumber 
\err_\MLE^2 & : = \frac{2.58^2}{n}
 \vy^T \left[ \mCInv - 
\frac{ \mCInv \vone \vone^T \mCInv }{\vone^T\mCInv \vone}
\right] \vy \\
\label{eqn:errMLE}
&\qquad \qquad \times 
(c_0 - \vc^T\mC^{-1}\vc ),
\end{align}
\begin{align}
\label{eqn_prob_CI_MLE}
\mathbb{P}_f \left[
|\mu-\hmu_\MLE| \leq \err_\MLE \right]  = 99\%.
\end{align}
Here $c_0$, $\vc$, and $\mC$ are assumed implicitly to be based on $\vtheta = \vtheta_\MLE$.   
\subsubsection{Full Bayes}
Rather than use maximum likelihood to determine $m$ and $s$ one can treat them as hyperparameters with a non-informative, conjugate prior, namely $\vrho_{m,s^2}(\xi, \lambda) \propto 1/\lambda$. Let us assume integral $\mu$, a random variable, taking typical values $z$, then the posterior density for the $\mu$ given data $\vy$ using Bayes theorem is
\begin{align*}
\MoveEqLeft[1]{\rho_{\mu}(z | \vf = \vy)} \\
& \propto \int_{0}^\infty \int_{-\infty}^\infty \rho_{\mu}(z | \vf = \vy, m = \xi, s^2 = \lambda) \\
& \hspace{2.5cm} \times  \rho_{\vf}(\vy | \xi, \lambda ) \rho_{m, s^2}(\xi, \lambda) \, \D \xi \D \lambda \\
& \propto \left( 1 +  \frac{(z - \hmu_{\textup{MLE}})^2}{(n-1)\widehat{\sigma}_{\textup{full}}^2} \right)^{-n/2}, \text{  (by Appendix \ref{appendix:full_bayes})}
\end{align*}
where 
\begin{multline*}
%\label{FJH:eq:signmaFull}
\hsigma^2_{\textup{full}} 
:= \frac{1}{n-1}
\vy^T\left[ \mC^{-1} 
- \frac{ \mC^{-1} \vone\vone^T \mC^{-1}}{\vone^T \mC^{-1} \vone}  \right]\vy
\\ 
\times  \left[\frac{(1 - \vc^T \mC^{-1} \vone)^2}{\vone^T \mC^{-1} \vone} + (c_0  -\vc ^T \mC^{-1} \vc) \right].
\end{multline*}

@@

Because the shape parameter, $\vtheta$, enters the definition of the covariance kernel in a non-trivial way, the only way to treat it as a hyperparameter and assign a tractable prior would be for the prior to be discrete.  We believe in practice that choosing such a prior involves more guesswork, so we choose to use either empirical Bayes estimate of $\vtheta$ in \eqref{eqn:thetaMLE} or cross-validation approach described next.

@@

The loss function $\textup{GCV}$ depends on $m$ and $\vtheta$, but not on $s$.  Minimizing the GCV  yields
\begin{align*}
m_{\textup{GCV}} = \frac{\vone^T \mC^{-2} \vy}{\vone^T \mC^{-2} \vone},
\end{align*}
\begin{multline}
\label{vthetaGCV}
\vtheta_{\textup{GCV}} = \argmin_\vtheta \biggl\{\log \left(  \vy^T \left[\mC^{-2} - \frac{\mC^{-2} \vone \vone^T \mC^{-2}}{\vone^T \mC^{-2} \vone}  \right] \vy \right)  \\
- 2 \log \left ( \trace(\mC^{-1}) \right ) \biggr\}.
\end{multline}
Plugging the value of $m_\GCV$ into \eqref{eqn:BayesCub} yields
\begin{align}
\label{eqn:muCV}
\widehat{\mu}_{\textup{GCV}}
= \left(\frac{(1 - \vone^T  \mC^{-1}\vc) \mC^{-1} \vone}{\vone^T \mC^{-2} \vone} + \vc \right)^T \mC^{-1} \vy.
\end{align}

@@

To demonstrate automatic Bayesian cubature consider a Mat\'ern covariance kernel:
\begin{align*}
C_{\theta}(\vx, \vt) = \prod_{k=1}^d \exp(-\theta|\vx_k-\vt_k|)(1+\theta |\vx_k-\vt_k|),
\end{align*}
and Sobol points as the nodes.
Also, consider the integration problem of evaluating  \emph{multivariate normal probabilities}:
\begin{align}
\label{eqn:GaussDef}
\mu = \int_{(\va,\vb)} \frac{\exp\bigl(- \frac 12 \vt^T \mSigma^{-1} \vt \bigr)}{\sqrt{(2 \pi)^{d'} \det(\mSigma)}} \, \dvt,
\end{align}
where $(\va,\vb)$ is a finite, semi-infinite or infinite box in $\reals^{d'}$.  This integral does not have an analytic expression for general $\mSigma$, so cubatures are required.

Genz \cite{Gen93} introduced a variable transformation to transform \eqref{eqn:GaussDef} into an integral on the unit cube.  Let $\mSigma= \mL \mL^T$ be the Cholesky decomposition where $\mL = (l_{jk})_{j,k=1}^d$ is a lower triangular matrix.  Iteratively define
\begin{align*}
\alpha_1& = \Phi(a_1), \qquad \beta_1 = \Phi(b_1)
\\
\alpha_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
a_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\\
\beta_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
b_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\end{align*}
\begin{align}
\label{fGenzdef}
f_{\text{Genz}}(\vx) = \prod_{j=1}^d [\beta_j(\vx) - \alpha_j(\vx)].
\end{align}
where $\Phi$ is the cumulative standard normal distribution function.  Then, $\mu = \int_{[0,1]^{d'-1}} f_{\text{Genz}}(\vx) \, \dvx$.

\begin{figure}
	\captionsetup[subfigure]{labelformat=empty}
	%\begin{subfigure}[h]{0.48\linewidth}
	%	\includegraphics[width=1.1\linewidth]{Plotting_gaussian}
	%\end{subfigure}
	\centering
	%\begin{subfigure}[h]{0.48\linewidth}
		\includegraphics[width=0.7\linewidth]{GenzFunc_varTx_none}
	%\end{subfigure}
	\caption{The $d'=3$ multivariate normal probability transformed to an integral of $f_{\text{Genz}}$ with  $d=2$. This plot can be reproduced using \code{IntegrandPlots.m} in GAIL}
	\label{fig:MVN_Genz}
\end{figure}

We use the following parameter values in the simulation: 
\begin{align*}
d' = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol points \cite{DicEtal14a,DicPil10a}.  The results for  $\varepsilon = 10^{-2}, 10^{-3}, 10^{-4}$, and $10^{-5}$ and $100$ scrambles for each $\varepsilon$ are shown in \figref{fig:MVN_Metern_d2b2}.  We  observe the algorithm meets the error criterion 95\% of the time.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer, 
As shown in \figref{fig:MVN_Metern_d2b2}, computation time increases rapidly with $n$. 
The  maximum likelihood estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

@@

\subsection{Fast Transform Kernel}
We make some assumptions about the relationship between the covariance kernel and the nodes, which will be shown to hold in Section \ref{sec:shift_invariant_kernel} for rank-1 lattices and shift-invariant kernels.  First we introduce the notation
\begin{align}
\nonumber
\mC &= \Big(C_\vtheta(\vx_i,\vx_j)\Big)_{i,j=1}^n  = (\vC_1,...,\vC_n) 
\\
\label{eqn:ftk_factor}
&= \frac 1n \mV \mLambda \mV^H , 
\quad \quad \mV^H = n \mV^{-1}, \\
\nonumber
\mV &= (\vv_1,...,\vv_n)^T = (\vV_1,...,\vV_n), \\
\nonumber
\mC^p  &= \frac 1n \mV \mLambda^{p} \mV^H, \qquad \forall p \in \integers,
\end{align}
where $\mV^H$ is the Hermitian of $\mV$, $\vC_1,...,\vC_n$ are columns of $\mC$.  $\vV_1,...,\vV_n$ are columns of $\mV$ and $\vv_1,...,\vv_n$ are rows of $\mV$.  The columns of matrix $\mV$ are eigenvectors of $\mC$, and $\mLambda$ is a diagonal matrix of eigenvalues of $\mC$.
For any $n \times n$ vector $\vb$, define the notation  $\widetilde{\vb} := \mV^H \vb$.

@@

Under assumptions \eqref{fastcompAssump} the eigenvalues may be identified as the fast transform of the first column of $\mC$:
\begin{align}
\nonumber
\vlambda 
& = \begin{pmatrix}
\lambda_1 \\ \vdots \\ \lambda_n
\end{pmatrix} = \mLambda \vone = \mLambda \vv_1^* 
= \underbrace{\left( \frac 1n \mV^H  \mV \right) }_{\mathsf{I}} \mLambda \vv_1^* \\
&= \mV^H \left( \frac 1n \mV \mLambda \vv_1^* \right)
= \mV^H \vC_1 =  \widetilde{\vC}_1,
\label{eqn:fast_transform_to_eigvalues}
\end{align}
Where $\mathsf{I}$ is the identity matrix.
Also note that the fast transform of $\vone$ has a simple form
\begin{align*} 
\widetilde{\vone}
& = \mV^H \vone = \mV^H \vV_1 = 
\left(n, 0, \dots, 0 \right)^T.
% \begin{pmatrix}n  \\ 0 \\ \vdots \\ 0 \end{pmatrix}.
\label{eqn:fast_transform_one}
\end{align*}
